Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity\vspace{-7ex}

### 2016-08-07 papers/kmeans/kmeans.tex
Fix teh typo; added a 'the'

There are a few points to highlight about the abundance anti-correlations
found in globular clusters. First, the detection of these signatures gives
confirmation to our analysis: The training (reference) set was not purposefully selected
to contain members of any globular or open cluster. However amongst the test
set, we recover globular cluster members and their anti-correlations 
in chemical abundances. That is to
say we recover peculiar chemical abundances that are not dominant in our
training set, but are expected from previous studies. Secondly, the very
presence of complex abundance correlations makes the detection of clusters
(by k-means) substantially more difficult. Briefly, the k-means algorithm
is most effective for near-normally distributed, isotropic clusters in high-dimensional space.
While globular clusters often show distributions of this kind in a few dimensions,
they also demonstrate non-linear correlations which would be
sub-optimally selected by k-means. For these reasons the fact that (a)~\thecannon\
reports these abundance correlations and (b)~k-means (as a simple algorithm
without information about expectations in abundance space) does not appear to be
severely impacted by these correlations, gives us great hope for more sophisticated approaches.
That said, globular cluster members are over-represented in teh \apogee\ data set,
because many of them were specifically targeted for observations for being cluster
members \citep{apogee}; that is, globular-cluster structure is over-represented relative to the
field in this data set.

@@

The k-means algorithm employed here is by no means optimal, and other clusters
we identify are accompanied by a few stars that are not currently co-spatial.
Those stars may be unassociated interlopers that have been misclassified,
or they may be true cluster members that are now unbound. However in the
case of the high-velocity-dispersion structure in \figurename~\ref{fig:halo},
the situation is far more extreme. We identify a group of stars with
very similar abundances in 15 dimensions that are now spread throughout
the Galaxy halo.
Because some stars ($<8\%$) are candidate members of the massive globular
cluster M53, it provides tantalizing possibility that these stars may
indeed be all members of the same co-natal gas cloud, implying that 
these stars have been accreted onto the halo from a single globular
cluster early in the Milky Way's formation. While speculative,
this idea demonstrates the promise of chemical tagging.

>>>>

There are a few points to highlight about the abundance anti-correlations
found in globular clusters. First, the detection of these signatures gives
confirmation to our analysis: The training (reference) set was not purposefully selected
to contain members of any globular or open cluster. However amongst the test
set, we recover globular cluster members and their anti-correlations 
in chemical abundances. That is to
say we recover peculiar chemical abundances that are not dominant in our
training set, but are expected from previous studies. Secondly, the very
presence of complex abundance correlations makes the detection of clusters
(by k-means) substantially more difficult. Briefly, the k-means algorithm
is most effective for near-normally distributed, isotropic clusters in high-dimensional space.
While globular clusters often show distributions of this kind in a few dimensions,
they also demonstrate non-linear correlations which would be
sub-optimally selected by k-means. For these reasons the fact that (a)~\thecannon\
reports these abundance correlations and (b)~k-means (as a simple algorithm
without information about expectations in abundance space) does not appear to be
severely impacted by these correlations, gives us great hope for more sophisticated approaches.
That said, globular cluster members are over-represented in the \apogee\ data set,
because many of them were specifically targeted for observations for being cluster
members \citep{apogee}; that is, globular-cluster structure is over-represented relative to the
field in this data set.

@@

The k-means algorithm employed here is by no means optimal, and other clusters
we identify are accompanied by a few stars that are not currently co-spatial.
Those stars may be unassociated interlopers that have been misclassified,
or they may be true cluster members that are now unbound. However in the
case of the high-velocity-dispersion structure in \figurename~\ref{fig:halo},
the situation is far more extreme. We identify a group of stars with
very similar abundances in 15 dimensions that are now spread throughout
the Galaxy halo.
Because some stars ($<8\%$) are candidate members of the massive globular
cluster M53, it provides the tantalizing possibility that these stars may
indeed be all members of the same co-natal gas cloud, implying that 
these stars have been accreted onto the halo from a single globular
cluster early in the Milky Way's formation. While speculative,
this idea demonstrates the promise of chemical tagging.
